### 2024-09-27 paper/paper.qmd
More edits on discussion

---
title: "Marriage Data analysis"
subtitle: "STA304 WEEK 3"
author: 
  - Maggie Zhang
thanks: "Code and data are available at:https://github.com/MaggieZ111119/CityofToronto_Daily_Shelter_Occupancy.Rproj.git"
date: today
date-format: long
abstract: "This paper investigates the occupancy of shelters in Toronto from 2017 to 2020, focusing on capacity and usage across different shelter sectors. By analyzing the data, we explore how shelter demand {exceeded available resources in certain sectors. This research highlights significant challenges faced by the shelter system, providing insights into the strain on resources and potential gaps in support for the homeless population}. These findings are crucial for informing policies and improving resource allocation to better address homelessness in Toronto."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# Preparation

#Load used libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(knitr)
library(gridExtra)

# Read Data(Quarto need data set to be readed in the file)
shelter_data <- read_csv(here::here("~/CityofToronto_Daily_Shelter_Occupancy.Rproj/data/analysis_data/full_shelter_data.csv"), show_col_types = FALSE)
```

# Introduction

Homelessness remains a persistent issue across Canada, including in Toronto(@gaetz2010struggle). s one of the country’s largest cities, Toronto faces ongoing challenges in assisting individuals who, primarily due to poverty, cannot afford housing and struggle to meet basic needs(@jadidzadeh2018patterns). The City of Toronto operates a network of shelters catering to men, women, youth, and families, making it essential to balance the availability of these services with demand. Monitoring shelter occupancy trends with capacity is crucial for identifying resource gaps and developing effective policies to address the needs of homeless populations.

This paper analyzes shelter occupancy data from 2017 to 2020 in the City of Toronto, sourced from the Open Data Toronto portal (@OepnDataToronto). By examining occupancy trends during this period, particularly during the COVID-19 pandemic, we aim to evaluate how effectively the city's shelters met the increased demand. Our focus on specific sectors within the shelter system will help determine which groups were disproportionately affected and shed light on the system's adaptability in response to the crisis, particularly in addressing issues of overcrowding in family shelters.

By studying the shelter occupancy data from 2017 to 2020 across four cities, key findings emerged. There was an increase in shelter occupancy from late 2019 to early 2020, which coincided with the onset of the COVID-19 pandemic, particularly in Scarborough, which was observed to have higher than usual capacities of sheltering service during that period. Despite the rise in shelter use, the utilization rates remained high, indicating that the shelters were effectively meeting demand without being overburdened. The government's timely increase in shelter capacity during the pandemic was especially evident. Over Capacity sheltering experience exists, in Toronto's shelters for families, but it likely resulted from flexible accommodation arrangements rather than systemic flaws. Overall, the shelter system demonstrated well adaptability and support for homeless populations during this period.

The remainder of this paper is presented in sections: @sec-data :Data, @sec-results :Results, @sec-discussion :Discussion, and @sec-weakness :Weaknesses and Next Steps. The data section introduces the dataset used for analysis, explaining its source and the context that shaped its collection. This section thoroughly summarizes the data using visualizations created with `ggplot2` (@R-ggplot2) and tables produced with `knitr` (@R-knitr), also with `gridExtra` (@gridExtra) to help formatting. Key variables chosen for the analysis, along with the reasoning behind their selection, are explained. The results section presents the findings drawn from the data, highlighting patterns in shelter occupancy and compare to its availability. Followed by the discussion section, which explores the implications of these findings, particularly in relation to resource allocation and policy development, offering insights into how Toronto's shelter system responded to increasing demand.

# Data {#sec-data}
All data reviewed and analysis in this paper is the Daily Shelter Occupancy @Shelter. The data is essential for evaluating how shelters serve different populations (e.g., men, women, youth, families) and for analyzing the current shelter servise system over time, such as capacity shortfalls or surpluses. It provides inforamtion on all the active shelters exist in the City of Toronto area, collected in forms of four separate datasets corresponding to the years 2017, 2018, 2019, and 2020.

## Overview {#sec-data-overview}
Each datasets include various characteristics of the shelters. The data was sourced from an Open Data Toronto Portal @OpenDataToronto, with each row indicate a unique entry for a specific sector of specific shelter, on a specific date, and has Unique row identifier for Open Data database, "_id".

The collection process records information about corresponding name of the non-profit entity that is responsible for the shelter, as well as the name of the facility (e.g. hotel, residence building), along with and their capacity to accommodate homeless clients. The Capacity is being measured as number of bed or a mat/cot available. Something will happen is that for the family sector, it is possible to exceed available capacity depending on the bed number, because family could be accommodated in a room with number beds smaller than number of people in their family.  The data was collected every day, 4:00 AM, to record number of homeless clients occupying the shelters. This way to gather information about occupancy provides a consistent snapshot of shelter use across all organizations.

Other similar datasets, especialy the ther dataset publish by Toronto Shelter & Support Services, were considered. However, they did not offer the cross-sector view of both occupancy and capacity needed for the focus of this analysis. The selected dataset allows for a broad comparison across time and sectors in different city, providing a fuller picture of trends within the shelter system.

## Feature Selection and Aggregation

For the purpose of this analysis, several features are chose to be particularly focused on. variable of interest including the the city (named "SHELTER_CITY" in the dataset) in which shelters are located; the date of data recorded("OCCUPANCY_DATE"); the sector("SECTOR") of clientele the shelters serve, for example men, women, youth, families, and even co-ed; the shelter's capacities("CAPACITY") and occupancy("OCCUPANCY") levels. These variables were chosen for their relevance in which is to investigate shelter occupancy and capacity across different cities and sectors, focusing on the City of Toronto from 2017 to 2020.

There are also seven other detailed feature in the datasets: "ORGANIZATION_NAME", "SHELTER_NAME", "SHELTER_ADDRESS", "SHELTER_PROVINCE", "SHELTER_POSTAL_CODE", "FACILITY_NAME", and "PROGRAM_NAME". Using R programming lanaguage @citeR, the `janitor`@R-janitor, `tidyverse` @R-tidyverse, and `dplyr` @R-dplyr pakages are used in data simulation, downloading, cleaning, and writting test. No new variables were created for this analysis, but the data was aggregated by grouping observations by city, date, and sector. This aggregation facilitated a more comprehensive comparison across cities and sector groups. A random sample of the cleaned data can be seen in @fig-shelter-sample.
```{r}
#| label: fig-shelter-sample
#| tbl-cap: Sample of Cleaned Shelter Data
#| echo: false
#| message: false

set.seed(1009633096)
indices <- sample(1:nrow(shelter_data), 5)

kable(shelter_data[indices,],
      caption = "Sample of Cleaned Shelter Data")

```

## Data Breakdown
As shown in @fig-shelter-full-table, The four cities being observed in the dataset are Toronto, Etobicoke, North York, and Scarborough. The sectors recorded include Co-ed (mixture), Families, Men, Women, and Youth. Across all cities, record for Youth age group being presents. In fact, Youth is the only population in Etobicoke and North York. In Toronto, which has the most number of observation recorded, equal number of sectors are being observed, but how these people in distributed inthe total popluation are not clear. This will be further addressed in the following sections.
```{r}
#| label: fig-shelter-full-table
#| tbl-cap: Number of Observations by City and Sector
#| echo: false
#| message: false

# Count the number of observations by city and sector
city_sector_breakdown <- shelter_data %>%
  group_by(shelter_city, sector) %>%
  count() %>%
  spread(key = sector, value = n, fill = 0)

# Check if the total is being calculated correctly
city_sector_breakdown$Sum <- rowSums(city_sector_breakdown[,2:6], na.rm = TRUE)

# Display the table with correct formatting
city_sector_breakdown %>%
  kable(caption = "Number of Observations by City and Sector")

```

The following table @fig-shelter-diff-city provides a summary of the average, maximum, and minimum occupancy, as well as capacity metrics for each city. Toronto has most amount of shelter avaliable and also has most people occuupaying in. The Maximum occupany of Toronto in a day, combined all sectors, from 2017 to 2020,is 6803. Wheras Etobicoke only has 53 people occupanying. The fact that Etotobike has only hoemless people in the group of Youth recorded could explain alittle bit. But, the de=ifference still strong enough to see that the homeless population might be significant larger in Toronto compared to aother cities. 
```{r}
#| label: fig-shelter-diff-city
#| tbl-cap: Statistics for Homeless Shelter Occupancy by City
#| echo: false
#| message: false

# Because the data originally has one date and is separated into several rows with different sectors. 
# If I want to calculate the total occupancy of that date in that city, I need to add all sector information together. Then, we will see the average, max, mix of occupancy. 
# Same idea to capacity.

toronto_data <- shelter_data %>% 
  filter(shelter_city == "Toronto") %>% 
  group_by(occupancy_date) %>%
  summarise(
    total_occupancy = sum(total_occupancy, na.rm = TRUE),
    total_capacity = sum(total_capacity, na.rm = TRUE),
    .groups = 'drop'
  )

north_york_data <- shelter_data %>% 
  filter(shelter_city == "North York") %>%
  group_by(occupancy_date) %>%
  summarise(
    total_occupancy = sum(total_occupancy, na.rm = TRUE),
    total_capacity = sum(total_capacity, na.rm = TRUE),
    .groups = 'drop'
  )

scarborough_data <- shelter_data %>% 
  filter(shelter_city == "Scarborough")%>%
  group_by(occupancy_date) %>%
  summarise(
    total_occupancy = sum(total_occupancy, na.rm = TRUE),
    total_capacity = sum(total_capacity, na.rm = TRUE),
    .groups = 'drop'
  )

etobicoke_data <- shelter_data %>% 
  filter(shelter_city == "Etobicoke") %>%
  group_by(occupancy_date) %>%
  summarise(
    total_occupancy = sum(total_occupancy, na.rm = TRUE),
    total_capacity = sum(total_capacity, na.rm = TRUE),
    .groups = 'drop'
  )

# Summary statistics for each city
toronto_summary <- toronto_data %>%
  summarise(
    avg_occupancy = mean(total_occupancy, na.rm = TRUE),
    max_occupancy = max(total_occupancy, na.rm = TRUE),
    min_occupancy = min(total_occupancy, na.rm = TRUE),
    avg_capacity = mean(total_capacity, na.rm = TRUE),
    max_capacity = max(total_capacity, na.rm = TRUE),
    min_capacity = min(total_capacity, na.rm = TRUE)
  )

north_york_summary <- north_york_data %>%
  summarise(
    avg_occupancy = mean(total_occupancy, na.rm = TRUE),
    max_occupancy = max(total_occupancy, na.rm = TRUE),
    min_occupancy = min(total_occupancy, na.rm = TRUE),
    avg_capacity = mean(total_capacity, na.rm = TRUE),
    max_capacity = max(total_capacity, na.rm = TRUE),
    min_capacity = min(total_capacity, na.rm = TRUE)
  )

scarborough_summary <- scarborough_data %>%
  summarise(
    avg_occupancy = mean(total_occupancy, na.rm = TRUE),
    max_occupancy = max(total_occupancy, na.rm = TRUE),
    min_occupancy = min(total_occupancy, na.rm = TRUE),
    avg_capacity = mean(total_capacity, na.rm = TRUE),
    max_capacity = max(total_capacity, na.rm = TRUE),
    min_capacity = min(total_capacity, na.rm = TRUE)
  )

etobicoke_summary <- etobicoke_data %>%
  summarise(
    avg_occupancy = mean(total_occupancy, na.rm = TRUE),
    max_occupancy = max(total_occupancy, na.rm = TRUE),
    min_occupancy = min(total_occupancy, na.rm = TRUE),
    avg_capacity = mean(total_capacity, na.rm = TRUE),
    max_capacity = max(total_capacity, na.rm = TRUE),
    min_capacity = min(total_capacity, na.rm = TRUE)
  )

city <- c("Toronto", "North York", "Scarborough", "Etobicoke")
combined_summary <- bind_rows(toronto_summary,
                          north_york_summary,
                          scarborough_summary,
                          etobicoke_summary)
combined_summary$city <- city
combined_summary <- combined_summary %>%
  select(city, everything())

kable(combined_summary, 
             col.names = c("City", 
                           "Average Occupancy", 
                           "Max Occupancy", 
                           "Min Occupancy", 
                           "Average Capacity", 
                           "Max Capacity", 
                           "Min Capacity"),
      caption = "Statistics for Homeless Shelter Occupancy by City")
```
The relastion ship of two key feature, occupancy and capacity, are essential for understaidng current shelter servise system. The relashipship is reflected in @fig-occupancy-capacity-relashionship. Occupany and capacity are greatly follwing a positive correlation, indicating that most there aren't much over-investment to the shelter system, nor heavily over-filled shelter. But there are noticible deviation from the perfect correlation to above, these are suggecting there are cases where occupany is higher than capacity, suggesting occupantee may living in a overly paked places. More will be dicussed in  @sec-results, including other learning on the shelter servise like thier trend, and how these shelter information possiblly reflect the homeless.
```{r}
#| label: fig-occupancy-capacity-relashionship
#| fig-cap: Relationship Between Total Occupancy and Total Capacity in Shelter of the City of Toronto
#| fig-pos: "H"
#| echo: false
#| message: false

ggplot(shelter_data, aes(x = total_capacity, y = total_occupancy, color = sector)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship Between Total Occupancy and Total Capacity in Shelter of the City of Toronto",
       x = "Total Capacity",
       y = "Total Occupancy",
       color = "Sector") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

```

# Results {#sec-results}
Through grpahfical study of the datset, some usful information are found and the result will be dicussed in this section.

## Shelter Occupancy Trend over Time
The figure @fig-shelter-over-time depicts the trend of homeless people occupying shelters in each city over time. A noticeable peak in occupancy is observed across all four cities between late 2019 and early 2020. Especially, there is an rising trend in Scarborough. This rise coincides with the outbreak of COVID-19, as noted in the "About Daily Shelter Occupancy" section in the data website @ShelterThe. significant decrease in occupancy from mid-2020 across all cities further supports this correlation, suggesting the pandemic had a direct impact on shelter use.
```{r}
#| label: fig-shelter-over-time
#| fig-cap: Visualization of Shelter Occupancy During Time Period 2017 to 2020
#| echo: false
#| message: false

# Assuming you have a data frame 'shelter_data' with total occupancy for each city
# Create individual plots for each city
toronto_plot <- ggplot(toronto_data %>% slice(seq(1, n(), by = 50)), aes(x = occupancy_date, y = total_occupancy)) +
  geom_line(color = "blue") +
  labs(title = "Toronto Occupancy Trends",
       x = "Date",
       y = "Total Occupancy") +
  theme_minimal()

north_york_plot <- ggplot(north_york_data%>% slice(seq(1, n(), by = 50)), aes(x = occupancy_date, y = total_occupancy)) +
  geom_line(color = "green") +
  labs(title = "North York Occupancy Trends",
       x = "Date",
       y = "Total Occupancy") +
  theme_minimal()

scarborough_plot <- ggplot(scarborough_data%>% slice(seq(1, n(), by = 50)), aes(x = occupancy_date, y = total_occupancy)) +
  geom_line(color = "red") +
  labs(title = "Scarborough Occupancy Trends",
       x = "Date",
       y = "Total Occupancy") +
  theme_minimal()

etobicoke_plot <- ggplot(etobicoke_data%>% slice(seq(1, n(), by = 50)), aes(x = occupancy_date, y = total_occupancy)) +
  geom_line(color = "orange") +
  labs(title = "Etobicoke Occupancy Trends",
       x = "Date",
       y = "Total Occupancy") +
  theme_minimal()

combined_plot <- grid.arrange(toronto_plot, north_york_plot, 
                               scarborough_plot, etobicoke_plot, 
                               nrow = 2, ncol = 2,
                               top = "Visualization of Shelter Occupancy During Time Period 2017 to 2020"
                          )

# This will suppress the grid output in the document
invisible(combined_plot)

```

## Shelter Utilization Analysis
The Utilization Rate as being calculated as:
$$\text{Utilization Rate} = \left( \frac{\text{Number of Occupancy}}{\text{Number of Capacity}} \right) \times 100\%$$
As shown in @fig-shelter-utilization, Toronto almost consistently has the lowest utilization rate compared to the other cities over four years, although the rates remain relatively high. This suggests that shelter services are well-used and not over-invested, meaning that the shelters are indeed serving the homeless population effectively. Despite an increase in occupancy around the end of 2019, utilization rates decreased during this period, likely due to an increase in available shelter spaces provided during the COVID-19 pandemic, suggested in @fig-shelter-over-time. This incicated that more shelter are timely provided, the goverment are providing larger amount of shelter than usual during pandemic. This is particularly evident in Scarborough, where a sharp increase in shelter capacity is visible at the end of 2019 and the start of 2020. Refers to @fig-shelter-provided to observe more capacity changes in shelters over time.
```{r}
#| label: fig-shelter-utilization
#| fig-cap: Yearly change in Utilization Rate Average on All Shelters by City (2017-2020)
#| echo: false
#| message: false
#| warning: false

# Calculate utilization rate by city
toronto_data <- toronto_data %>%
  mutate(utilization_rate = total_occupancy / total_capacity)

north_york_data <- north_york_data %>%
  mutate(utilization_rate = total_occupancy / total_capacity)

scarborough_data <- scarborough_data %>%
  mutate(utilization_rate = total_occupancy / total_capacity)

etobicoke_data <- etobicoke_data %>%
  mutate(utilization_rate = total_occupancy / total_capacity)

# Combine all city datasets
combined_data <- bind_rows(
  toronto_data %>% mutate(shelter_city = "Toronto"),
  north_york_data %>% mutate(shelter_city = "North York"),
  scarborough_data %>% mutate(shelter_city = "Scarborough"),
  etobicoke_data %>% mutate(shelter_city = "Etobicoke")
)

# Extract year from occupancy_date
combined_data <- combined_data %>%
  mutate(year = year(occupancy_date),
         month = month(occupancy_date),
         season = case_when(
           month %in% c(12, 1, 2) ~ "Winter",
           month %in% c(3, 4, 5) ~ "Spring",
           month %in% c(6, 7, 8) ~ "Summer",
           month %in% c(9, 10, 11) ~ "Fall"
         ))

# Have data into seasons
combined_data <- combined_data %>%
  mutate(season_year = factor(paste(season, year), 
                               levels = c("Winter 2017", "Spring 2017", "Summer 2017", "Fall 2017",
                                          "Winter 2018", "Spring 2018", "Summer 2018", "Fall 2018",
                                          "Winter 2019", "Spring 2019", "Summer 2019", "Fall 2019",
                                          "Winter 2020", "Spring 2020", "Summer 2020", "Fall 2020")))

# Calculate average utilization rate
annual_utilization <- combined_data %>%
  group_by(season_year, shelter_city) %>%
  summarise(avg_utilization_rate = mean(utilization_rate, na.rm = TRUE), .groups = 'drop')

ggplot(annual_utilization, aes(x = season_year, y = avg_utilization_rate, color = shelter_city, group = shelter_city)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(x = "Seasons",
       y = "Average Utilization Rate",
       color = "City",
       caption = "Utilization Rate Average Change on All Shelters by City") +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format(scale = 100)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
#| label: fig-shelter-provided
#| fig-cap: Shelter Capacity During Time Period 2017 to 2020 of CIties
#| echo: false
#| message: false
#| warning: false

# Create individual plots for each city
toronto_plot <- ggplot(
  toronto_data %>% slice(seq(1, n(), by = 100)), 
                         aes(x = occupancy_date, y = total_capacity)) +
  geom_line(size = 1, color = "blue") +
  geom_point(size = 2) +
  labs(title = "Toronto", x = "Year", y = "Capacity") +
  theme_minimal()

north_york_plot <- ggplot(
  north_york_data %>% slice(seq(1, n(), by = 100)), 
  aes(x = occupancy_date, y = total_capacity)) +
  geom_line(size = 1, color = "green") +
  geom_point(size = 2) +
  labs(title = "North York", x = "Year", y = "Capacity") +
  theme_minimal()

scarborough_plot <- ggplot(
  scarborough_data %>% slice(seq(1, n(), by = 100)), 
  aes(x = occupancy_date, y = total_capacity)) +
  geom_line(size = 1, color = "red") +
  geom_point(size = 2) +
  labs(title = "Scarborough", x = "Year", y = "Capacity") +
  theme_minimal()

etobicoke_plot <- ggplot(
  etobicoke_data %>% slice(seq(1, n(), by = 100)), 
  aes(x = occupancy_date, y = total_capacity)) +
  geom_line(size = 1, color = "orange") +
  geom_point(size = 2) +
  labs(title = "Etobicoke", x = "Year", y = "Capacity") +
  theme_minimal()

# Combine the four plots into a single figure
combined_plot_cap <- grid.arrange(toronto_plot, north_york_plot, 
                              scarborough_plot, etobicoke_plot, 
                              nrow = 2, ncol = 2,
                              top = "Visualization of Shelter Capacity During Time Period 2017 to 2020")

# Display the combined plot
invisible(combined_plot_cap)

```

## Over Capacity Cases
The relationship between occupancy and capacity over the four years (@fig-occupancy-capacity-relashionship) reveals instances where shelter capacity was exceeded, as we discussed in the Data section (@sec-data). These cases where capacity is exceeded (when $occupancy > capacity$) are important to study when considering improvement shelter system improvement. By further investigation of the data, these overcrowded cases are primarily observed in the Families sector in Toronto (see @fig-overcapacity). Other sectors did not experience overcapacity, indicating that the shelter services were generally adequate. The possible overcrowding experience that happens to the Families sector may not necessarily reflect poor shelter conditions, as it could result from families being assigned to rooms with slightly fewer beds than their family size would suggest. For example, children might share beds (@Shelter), which does not indicate a systemic issue in shelter management.

```{r}
#| label: fig-overcapacity
#| tbl-cap: Overcrowded Cases by City and Sector
#| echo: false
#| message: false

# Filter for overcrowded data
overcrowded_data <- shelter_data %>%
  filter(total_occupancy > total_capacity)

# Create a frequency table including all cities and sectors
frequency_table <- overcrowded_data %>%
  group_by(shelter_city, sector) %>%
  summarise(count = n(), .groups = 'drop') %>%
  complete(shelter_city, sector, fill = list(count = 0)) %>%
  pivot_wider(names_from = sector, values_from = count)

# Create a complete dataset of all combinations of cities and sectors
all_combinations <- expand.grid(
  shelter_city = unique(shelter_data$shelter_city),
  sector = unique(shelter_data$sector)
)

# Create a frequency table including all cities and sectors
frequency_table <- overcrowded_data %>%
  group_by(shelter_city, sector) %>%
  summarise(count = n(), .groups = 'drop') %>%
  right_join(all_combinations, by = c("shelter_city", "sector")) %>%
  replace_na(list(count = 0)) %>%
  pivot_wider(names_from = sector, values_from = count, values_fill = 0)

# View the frequency table with knitr
frequency_table %>%
  kable(caption = "Overcrowded Cases by City and Sector")

```

# Discussion {#sec-discussion}

## Shelter servise evaluation

Based on the findings, particularly from the utilization rate data @fig-shelter-utilization, the shelter service system in the City of Toronto appears to be well-structured. Utilization rates consistently fall within the 90%-100% range, indicating that shelters are effectively supporting the homeless population.

## COVID-19 Impact

The occupancy of shelters reached its peak around winter 2019 to early 2020, as seen in @fig-shelter-over-time. This increase is likely linked to the COVID-19 pandemic, as shelters saw heightened demand. After this peak, occupancy dropped significantly. Suggesting the efforts by local governments to provide additional resources during the pandemic. The government intention to support homeless population is presented by higher capacities of living resources provided (see @fig-shelter-provided). These might be through higher capacities of existing shelters or adding in new shelters.

## Weaknesses and next steps {#sec-weakness}

COVID-19 has significantly impacted homeless populations, and the effort to meet their needs continues even five years after the outbreak. This data provides evidence of government support during the pandemic, but it is not enough to evaluate the effectiveness of shelter services and other government support for people living in homelessness. More rigorous research is required. The decreased utilization rate during the COVID outbreak may indicate that more individuals are experiencing severe challenges, preventing them from accessing shelters. Furthermore, it is insufficient to comment on the current state of government support since this data only goes up to 2020.

Understanding the development and importance of support services among people experiencing homelessness is essential. Moreover, a critical issue is how this support should be delivered. It is important to provide assistance without disrupting individuals' lives, for instance, by avoiding practices that displace or isolate them from their communities, such as relocating them to shelters far from their usual support systems (@boucher2022they). This is a long journey to go.

# References